---
title: "DATA_605_Assignment_3_Thonn"
file: DThonn_Assign3_DATA_605
name: Daniel Thonn
output:
  html_document: default
  pdf_document: default
course: DATA 605 CUNY MSDA program
---

** DATA_605_Assignment_3_Thonn - Eigenvalues & Eigenvectors **

http://rpubs.com/danthonn/361569

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r, echo=TRUE}
#install.packages("matrixcalc")

```

** Problem Set 1 **

1-1). What is the rank of matrix A?

$A = \left[\begin{array}{rrrr} 1 & 2 & 3 & 4 \\ -1 & 0 & 1 & 3 \\ 0 & 1 & -2 & 1 \\ 5 & 4 & -2 & -3 \end{array} \right]$ 


``````{r, echo=TRUE}

library(matrixcalc)

#create Matrix M1
A <- matrix(data = c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4, ncol = 4, byrow = TRUE)

A

matrix.rank(A)

# Answer 1-1: the rank of A is 4

# [1] 4

```


1-2).	Given an $m x n$ matrix where $m > n$, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero ? Maximum rank = $m$ (rows) Minimum rank = 1 (all other rows could be linearly dependent)

Answer-1-2:

Rank:

The maximum number of linearly independent vectors in a matrix is equal to the number of non-zero rows in its row echelon matrix. Therefore, to find the rank of a matrix, we transform the matrix to its row echelon form and count the number of non-zero independent rows.

View an example $m x n$ matrix
$$\left[ {\begin{array}{rrrr} a & b & c\\ d & e & f\\ g & h & i\\ j & k & l\ \end{array} } \right]$$


View in reduced form:
$$\left[ {\begin{array}{rrrr} x1 & 0 & 0\\ 0 & x2 & 0\\ 0 & 0 & x3\\ 0 & 0 & x4\\ \end{array} } \right]$$

Notice that the rows below the nth column become multiples of each other, and therefore can be eliminated.

Therefor the minimum rank of an m x n matrix that is non-zero is 1. 

The minimum rank is 1

And the maximimum rank is m (where m is the number of rows)


1-3). What is the rank of matrix $B$?

Rank:

The maximum number of linearly independent vectors in a matrix is equal to the number of non-zero rows in its row echelon matrix. Therefore, to find the rank of a matrix, we transform the matrix to its row echelon form and count the number of non-zero independent rows.


$$ B = \begin{bmatrix}1 & 2 & 1 \\3 & 6 & 3 \\2 & 4 & 2 \end{bmatrix}$$

row 2 is a multiple of row 1.  

$$ B = \begin{bmatrix}3 & 6 & 3 \\3 & 6 & 3 \\2 & 4 & 2 \end{bmatrix}$$ 
subtract row 1 from row 2.

$$ B = \begin{bmatrix}1 & 2 & 1 \\0 & 0 & 0 \\2 & 4 & 2 \end{bmatrix}$$ 
row 3 is a multiple of row 1.
$$ B = \begin{bmatrix}2 & 4 & 2 \\0 & 0 & 0 \\2 & 4 & 2 \end{bmatrix}$$ 
subtract row 1 from row 3.
$$ B = \begin{bmatrix}1 & 2 & 1 \\0 & 0 & 0 \\0 & 0 & 0 \end{bmatrix}$$ 
There is one independent row in this matrix.
$$ Rank =1 $$


1-3). What is the rank of matrix B?
``````{r, echo=TRUE}
B <- matrix(data = c(1,2,1,3,6,3,2,4,2), nrow = 3, ncol = 3, byrow = TRUE)
B

matrix.rank(B)

# Answer 1-3: the rank of A is 1

# [1] 1

```

** Problem Set 2 **

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

Eigenvalues:

$$A = \begin{pmatrix} 1 & 2 & 3\\ 0 & 4 & 5 \\ 0 & 0 & 6 \end{pmatrix}$$

$$ \lambda,I_3 = \begin{bmatrix}\lambda & 0 & 0 \\0 & \lambda & 0 \\0 & 0 & \lambda \end{bmatrix}$$ 


$$\lambda I_3 - A = \begin{pmatrix} \lambda -1 & -2 & -3\\ 0 & \lambda - 4 & -5 \\ 0 & 0 & \lambda -6 \end{pmatrix}$$

$$det ( \lambda I_3 -A ) = 0$$
$$(\lambda-6)(\lambda-4)(\lambda-1)-(-2)(0) +-3(0)$$
$$(\lambda-6)(\lambda-4)(\lambda-1) = 0$$

$$\lambda = 6 , \lambda=4 , \lambda = 1$$

characteristic polynomial = $$x^3 - 9x^2 +24x - 16$$

Eigenvectors:

1). lambda = 1
$$\lambda=1$$ 

$$ \begin{bmatrix}\lambda - 1 & 2 & 3 \\0 & \lambda - 4 & 5 \\0 & 0 & \lambda - 6 \end{bmatrix}$$

$$ \begin{bmatrix}0 & 2 & 3 \\0 & -3 & 5\\0 & 0 & -5\end{bmatrix},\begin{bmatrix}v_1 \\v_2 \\v_3\end{bmatrix}=0$$





$$first, pivot = 0,x_1 = free,with, value=1$$

$$-3 x_2 + 5 x_3 = 0 ,and, 5 x_3 = 0$$

$$x_{\lambda=1},=\begin{bmatrix}1 \\0 \\0\end{bmatrix}$$


2). lambda = 4
$$ \begin{bmatrix}\lambda - 1 & 2 & 3 \\0 & \lambda - 4 & 5 \\0 & 0 & \lambda - 6 \end{bmatrix}$$

$$\lambda=4$$ $$ \begin{bmatrix}3 & 2 & 3 \\0 & 0 & 5\\0 & 0 & -2\end{bmatrix},\begin{bmatrix}v_1 \\v_2 \\v_3\end{bmatrix}=0$$ 



$$Second,pivot = 0,x_2=free, with value=1.$$ 

$$3x_1+2x_2 +3x_3 = 0,and, 2x_3 = 0$$ 

$$x_3=0,x_2=1,and,x_1=-2/3$$ 




$$x_{\lambda=4},=\begin{bmatrix}-2/3\\1 \\0\end{bmatrix}$$

3). lambda = 6

$$ \begin{bmatrix}\lambda - 1 & 2 & 3 \\0 & \lambda - 4 & 5 \\0 & 0 & \lambda - 6 \end{bmatrix}$$


$$\lambda=6$$ $$ \begin{bmatrix}5 & 2 & 3\\0 & 2 & 5\\0 & 0 & 0\end{bmatrix},\begin{bmatrix}v_1 \\v_2 \\v_3\end{bmatrix}=0$$


$$Third,pivot =  0 ,x_3=free,with value=1.$$ 

$$5x_1 +2x_2 +3x_3 = 0 ,and, 2x_1+5x_3 = 0$$ 

$$x_3 = 1,x_2 = 2/5,and,x_1=-5/2$$

$$x_{\lambda=6},=\begin{bmatrix}-5/2 \\2/5 \\1\end{bmatrix}$$




``````{r, echo=TRUE}
A <- matrix(data = c(1,0,0,2,4,0,3,5,6), nrow = 3, ncol = 3, byrow = FALSE)
A

EV1 <- eigen(A)

EV1

EV2 <- EV1$values

EV2


```


**END**
